### 2024-08-15 parts/correction.tex
enormous rewrite

\AP
The counter example provided by \cref{def:bad-polynomial} relies on the fact
that $\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-polyregular functions} are. 
Indeed, the polynomial $\BadPoly$ introduced in
\cref{def:bad-polynomial} satisfies $\BadPoly(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative
coefficient for a \kl{maximal monomial}, namely $-2XY$.
In this section, we prove that closing
$\CoveredPoly$ under variable assignments is enough to recover from
\cref{karh:thm}.
We use the following notation to fix the value of some indeterminate, if
$P(X,Y)$ is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is
the polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial
function from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial
\Nat$, the restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with
indeterminates $\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the
variables of the domain of $\nu$.

\begin{definition}
    Let $\vec{X}$ be a finite tuple of indeterminates.
    The class $\intro*\CorrectPoly[\vec{X}]$ is the collection of
    polynomials $P \in \Rel[\vec{X}]$ such that,
    for every partial function $\nu \colon \vec{X} \topartial \Nat$,
    every \kl{maximal monomial} of
    $\restr{P}{\nu}$ is \kl{non-negative}.
\end{definition}

First, let us remark that $\CorrectPoly \subseteq \CoveredPoly$, because
polynomials in $\CorrectPoly$ are \kl{non-negative}. We also remarked at the
beginning of this section that our counter example provided in
\cref{thm:counter-example} is not in $\CorrectPoly$.

\AP Let us now prove that \kl{$\Nat$-rational polynomials} are in
$\CorrectPoly$. This follows from the correct implication in the statement of
\cref{karh:thm}, but we provide a self-contained proof using a refinement of
the classical combinatorial arguments for $\ZPoly$ \cite[Lemma 4.16]{CDTL23}
and $\NPoly$ \cite[Lemma 5.37]{DOUE23}. We believe that the formulation of such
pumping arguments can be made clearer through some extra notations, with the
added property of highlighting the central role of \kl{commutative} polyregular
functions. Let us define that a function $q$ is a \intro{(noncommutative output)
polynomial function} from $\Nat^p$ to $\Sigma^*$ whenever there exists words
$\alpha_0, \dots, \alpha_p \in \Sigma^*$, and words $u_1, \dots, u_p \in
\Sigma^*$, such that $q(X_1, \dots, X_p) = \alpha_0 \prod_{i = 1}^p u_i^{X_i}
\alpha_i$. That is, $q$ is syntactically defined by a non-commutative
polynomial function over the semi-ring $\Sigma^*$. Almost by definition,
\kl{polynomial functions} are \kl{commutative} \kl{polyregular functions}.

\AP There is still a minor detail that remains to be smoothened before proving
our combinatorial \cref{n-poly-combinatorics:lem}, which is that some
polyregular functions can exhibit a limited form of division. For instance, the
function computing the sum of the lengths of the suffixes of a word $w$
actually computes $\card{w} \times (\card{w} - 1) / 2$. This will be at odds
with our restriction to polynomials with coefficients in $\Rel$. However, we
argue that up to a suitable change of basis for the ring of polynomials, all
coefficients will actually be in $\Nat$. This relatively simple trick works by
considering as ``basic polynomials" the \intro{binomial monomial} of the form
$\binom{X - \ell}{k}$, where $\ell$ and $k$ are natural numbers. Beware that
$\binom{X - \ell}{k}$ is defined to be $0$ when $X \leq \ell$, and is therefore
a ``true polynomial" only when $X \geq \ell$. This is why we introduce the
notion of \intro{natural binomial polynomials} as $\Nat$-linear combinations of
products of \kl{binomial monomials}.

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be an \kl{$\Nat$-polyregular} function. 
    There exists a computable $\omega \in \Nat_{\geq 1}$
    such that for all \kl{polynomial functions}
    $q \colon \Nat^p \to \Sigma^*$,
    there exists a computable \kl{natural binomial polynomial} $P$
    such that:
    \begin{equation*}
        f \circ q(\omega X_1, \dots, \omega X_p)
        =
        P
        \quad 
        \text{ over } (\Nat_{\geq 1})^p
    \end{equation*}
\end{lemma}
\begin{proof}
    Let us fix a finite monoid $M$, a morphism $\mu \colon \Sigma^* \to M$, and 
    let $\omega$ be an idempotent power for the finite monoid $M$,
    i.e., a number such that for all $x \in M$,
    $x^{\omega+1} = x^\omega$.
    We will prove by induction on $k \in \Nat_{\geq 1}$
    that for all
    $\pi \colon M^k \to \Nat$,
    the lemma holds for the function 
    $f = \pi^\dagger$, i.e., such that
    $f(w)$ is the sum over all factorizations of $w$
    into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
    of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.
    Since every \kl{$\Nat$-polyregular function}
    is obtained via some choice of $M, \mu, \pi$ (\cref{nat-rel-poly:def})
    we derive the desired result.

@@

    \textbf{Base case: $k = 1$.}
    Let us show that $f \circ q$ is
    actually a constant (positive) function (and in particular a \kl{natural binomial polynomial}), as for
    all $n_1, \dots, n_p \geq 1$,
    \begin{align*}
        f(q(n_1, \dots, n_p))
        &\eqdef \pi\left(\mu(q(n_1, \dots, n_p))\right) \\
        &\eqdef \pi\left(\mu(\alpha_0 \prod_{i = 1}^p u_i^{\omega n_i} \alpha_i)\right)
        \\
        &= \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i)^{\omega n_i} \mu(\alpha_i)\right) 
        & \text{ morphism }
        \\
        &= \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i) \mu(\alpha_i)\right)
        & \text{ idempotent and } n_i \geq 1
        &= c \geq 0
    \end{align*}

@@

    We will use the fact that given $n_1, \dots, n_p \geq 0$,
    there exists a bijection between
    partitions of $q(\omega n_1, \dots, \omega n_p)$
    into $k+1$ words, and 
    triples composed of
    a position  $1 \leq \ell \leq p$,
    a ratio     $0 \leq b \leq n_\ell$,
    a remainder $0 \leq r < \omega \times \card{u_\ell}$,
    and a partition of
    $u_\ell^\omega\mid_{[-r:]} u_\ell^{\omega \times (n_\ell - b)}
    \alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{n_p} \alpha_{p}$
    into $k$ words,
    where we have used the notation 
    $v\mid_{[-r:]}$ to denote the word obtained from $v$ 
    by keeping only the last $r$ letters, in a sort of Pythonic syntax.
    That is, we look at where the first ``cut" is made.

    Now, given $(\ell, b, r)$ respecting the inequalities 
    mentioned above,
    one can define the \kl{polynomial functions}
    $s_{\ell,b,r} \colon \Nat^{\ell - 1} \to \Sigma^*$
    and
    $q_{\ell,b,r} \colon \Nat^{p - \ell + 1} \to \Sigma^*$
    so that
    $q(X_1, \dots, X_p) = s_{\ell, b, r}(X_1, \dots, X_{\ell-1}) 
    q_{\ell, b, r}(X_{\ell}, \dots, X_p)$
    under the assumption that $X_1 \geq b$:
    \begin{equation*}
        q_{\ell,b,r}(X_{1}, \dots, X_{p - \ell + 1})
        \defined
        u_\ell^\omega\mid_{[-r:]} u_\ell^{\omega \times (X_1 - b)}
        \alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{X_{p - \ell + 1}} \alpha_{p}
        \quad .
    \end{equation*}
    Similarly, let us define 
    $s_{\ell, b, r}(X_1, \dots, X_{\ell - 1})$
    as 
    $\alpha_0 u_1^{\omega X_1} \alpha_2 \dots \alpha_{\ell - 1} 
    u_\ell^{\omega b} u_{\ell}^\omega \mid_{[1:r]}$,
    Remark that $q_{\ell,b,r}$ has all of its indeterminates
    multiplied by $\omega$, so that the induction hypothesis 
    can be used.

    The key remark is that $\mu(s_{\ell,b,r}(X_1, \dots, X_{\ell - 1}))$ is a
    constant function from $\Nat^{\ell-1}$ to $M$, and that its value is also
    constant when $b$ varies among numbers greater than $0$. This holds with
    the exact same proof as the base case of this induction. Let us call
    $m_{\ell,r,1}$ and $m_{\ell,r,0}$ the two elements of the monoid that are
    obtained by $\mu(s_{\ell,b,r}(X_1, \dots, X_{\ell-1}))$ respectively when
    $b = 0$ and when $b > 0$. Finally, let us define $\pi_{\ell,r,0} \colon M^k
    \to \Nat$ via $\pi_{\ell,r,0}(m_2, \dots, m_{k+1}) \defined
    \pi(m_{\ell,r,0}, m_1, \dots, m_{k+1})$ and similarly for $\pi_{\ell,r,1}
    \colon M^k \to \Nat$.

    With all of this preliminary work done, we are now ready to conclude by
    simply unrolling the definition of 
    a function
    $f$ defined using a \kl{production function}
    $\pi \colon M^{k+1} \to \Nat$.
    In the following equations, we omit the precise
    range for the sum over $(\ell, b, r)$ to save space.
    Similarly, we will use 
    $q_{\ell,r}(X_1, \dots, X_{p-\ell+1})
    \defined q_{\ell,b,r}(X_1+b, X_2, \dots, X_{p - \ell + 1})$
    at the end of the computation.
    \begin{align*}
        &f(q(n_1, \dots, n_p))\\
        &\eqdef 
        \sum_{v_1 \dots v_{k+1} = q(n_1, \dots, n_p)}
        \pi(\mu(v_1), \dots, \mu(v_{k+1})) \\
        &= \sum_{(\ell, b, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,b,r}(n_\ell, \dots, n_p)}
           \pi(\mu(s_{\ell,b,r}(n_1, \dots, n_{\ell-1})), \mu(v_2), \dots, \mu(v_{k+1}))
        \\
        &= 
           \sum_{(\ell, 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,0,r}(n_\ell, \dots, n_p)}
           \pi(m_{\ell,r,0}, \mu(v_2), \dots, \mu(v_{k+1}))
           \\
        &+ 
           \sum_{(\ell, b > 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,b,r}(n_\ell, \dots, n_p)}
           \pi(m_{\ell,r,1}, \mu(v_2), \dots, \mu(v_{k+1}))
        \\
        &= 
           \sum_{(\ell, 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,0,r}(n_\ell, \dots, n_p)}
           \pi_{\ell,r,0}(\mu(v_2), \dots, \mu(v_{k+1}))
           \\
        &+ 
           \sum_{(\ell, b > 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,b,r}(n_\ell, \dots, n_p)}
           \pi_{\ell,r,1}(\mu(v_2), \dots, \mu(v_{k+1}))
        \\
        &= 
           \sum_{1 \leq \ell \leq p}
           \sum_{0 \leq r < \omega \times \card{u_\ell}}
           (\pi_{\ell,r,0})^\dagger (q_{\ell, 0, r}(n_\ell, \dots, n_p))
           \\
        &+ 
           \sum_{1 \leq \ell \leq p}
           \sum_{0 \leq r < \omega \times \card{u_\ell}}
           \sum_{0 < b < n_\ell}
           (\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
    \end{align*}
    Now, using the induction hypothesis,
    we have \kl{natural binomial polynomials} $P_{\ell,r,0}$ and $P_{\ell,r,1}$
    satisfying 
    \begin{equation*}
       (\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
       = P_{\ell,r,1}(\omega (n_\ell - b), \dots, \omega n_p)
       \quad 
       \forall n_\ell - b, \dots, n_p \geq 1
       \quad .
   \end{equation*}
   And similarly
    \begin{equation*}
        (\pi_{\ell,r,0})^{\dagger}(q_{\ell,r}(n_1, \dots, n_{\ell-1})
        = P_{\ell,r,0}(\omega n_1, \dots, \omega n_{\ell-1})
       \quad 
       \forall n_1, \dots, n_{\ell - 1} \geq 1
       \quad .
   \end{equation*}

   It is an easy and folklore fact that \kl{natural binomial polynomials} are
   stable under the operation $\Sigma_Y \colon Q(X_1,\dots,X_n,Y) \mapsto \sum_{0
   < b < Y} Q(X_1, \dots, X_n, b)$, and by pre-composition with 
   a multiplication by a constant $\omega \geq 1$.

   This allows us to conclude by defining 
   \begin{equation*}
       P(X_1, \dots, X_p) \defined
       \sum_{1 \leq \ell \leq p} \sum_{0 \leq r < \omega \card{u_\ell}}
       P_{l,r,0}(\omega X_{\ell}, \dots, \omega X_p) + \Sigma_Y P_{l,r,1}(\omega Y,
       \omega X_{\ell+1}, \dots, \omega X_p)
       \quad .
   \end{equation*}
   The latter being a \kl{natural binomial polynomial} satisfying
   the desired equality.
\end{proof}

Let us highlight that the multiplicative factor $\omega$ is necessary in
\cref{n-poly-combinatorics:lem}. Indeed, the function $f \colon \set{a}^* \to
\Nat$ defined as $0$ when the input is of odd length and $1$ when the input is
of even length is \kl{$\Nat$-polyregular}, but $f(a^X)$ is not a polynomial.
Let us remark that the \kl{maximal monomials} of the polynomial $\binom{X -
p}{k}$ have coefficient $1/k!$, and are therefore \kl{non-negative}.

\begin{corollary}
    \label{n-rat-correct:lem}
    Let $P \in \Rel[X_1, \dots, X_p]$ be an \kl{$\Nat$-rational polynomial}.
    Then,
    $P \in \CorrectPoly$.
    \proofref{n-rat-correct:lem}
\end{corollary}
\begin{proof}
    Let $f$ be the \kl{commutative}
    \kl{$\Nat$-polyregular function}
    with domain defined as $\Sigma \defined \set{a_1, \dots, a_p}$
    that \kl{represents} $P$. 
    Using \cref{n-poly-combinatorics:lem},
    there exists a number $\omega \in \Nat_{\geq 1}$
    and \kl{natural binomial polynomial} $Q$
    such that
    for all $n_1, \dots, n_p \geq 1$:
    \begin{equation*}
        f\left(
            \prod_{i = 1}^p a_i^{\omega n_i}
        \right)
        = Q(n_1, \dots, n_p)
        = P(\omega n_1, \dots, \omega n_p) 
        \quad .
    \end{equation*}
    We conclude that $P(\omega X_1, \dots, \omega X_p) = Q(X_1, \dots, X_p)$
    as polynomials,
    and in particular that
    the \kl{maximal monomials} of 
    $P$ are \kl{non-negative}.

>>>>

\AP
The counterexample provided by \cref{def:bad-polynomial} relies on the fact
that $\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-polyregular functions} are. 
Indeed, the polynomial $\BadPoly$ introduced in
\cref{def:bad-polynomial} satisfies $\BadPoly(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative
coefficient for a \kl{maximal monomial}, namely $-2XY$.
In this section, we prove that closing
$\CoveredPoly$ under variable assignments is enough to recover from
\cref{karh:thm}.
We use the following notation to fix the value of some indeterminate, if
$P(X,Y)$ is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is
the polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial
function from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial
\Nat$, the restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with
indeterminates $\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the
variables of the domain of $\nu$.

\begin{definition}
    The class $\intro*\CorrectPoly[\vec{X}]$ is the collection of
    polynomials $P \in \Rel[\vec{X}]$ such that,
    for every partial function $\nu \colon \vec{X} \topartial \Nat$,
    every \kl{maximal monomial} of
    $\restr{P}{\nu}$ is \kl{non-negative}.
\end{definition}

First, let us remark that $\CorrectPoly \subseteq \CoveredPoly$, because
polynomials in $\CorrectPoly$ are \kl{non-negative}. We also remarked at the
beginning of this section that our counterexample provided in
\cref{thm:counter-example} is not in $\CorrectPoly$.

\AP Let us now prove that \kl{$\Nat$-rational polynomials} are in
$\CorrectPoly$. This follows from the correct implication in the statement of
\cref{karh:thm}, but we provide a self-contained proof using a refinement of
the classical combinatorial arguments for $\ZPoly$ \cite[Lemma 4.16]{CDTL23}
and $\NPoly$ \cite[Lemma 5.37]{DOUE23}. We take extra care to reprove in
\cref{n-poly-combinatorics:lem} a finer statement that has two main goals:
first, it formulates the traditional \emph{pumping argument} as a composition
property using \kl{(noncommutative output) polynomial functions}, allowing us
to use the same formalism in \cref{k-combinatorial:def} and
\cref{ultimately-polynomial:def} and highlight the role of \kl{commutative}
\kl{polyregular functions} in the broader study of \kl{polyregular functions};
then we also give a precise shape of the polynomials that arise from such
\emph{pumping arguments}, which was lacking in former similar statements where
$\Rat[X_1, \dots, X_n]$ was used as an over approximation.

\AP To address our first goal, let us define that a function $q$ is a
\intro{(noncommutative output) polynomial function} from $\Nat^p$ to $\Sigma^*$
whenever there exists words $\alpha_0, \dots, \alpha_p \in \Sigma^*$, and words
$u_1, \dots, u_p \in \Sigma^*$, such that $q(X_1, \dots, X_p) = \alpha_0
\prod_{i = 1}^p u_i^{X_i} \alpha_i$. That is, $q$ is syntactically defined by a
non-commutative monomial over the semi- $\Sigma^*$. Almost by
definition, \kl{polynomial functions} are \kl{commutative} \kl{polyregular
functions}.

\AP Remark that our second goal is at odds with our restriction to polynomials
with coefficients in $\Rel$, because of the polynomial $X(X-1)/2$
(\cref{q-polynomial-n-poly:ex}). However, we argue that up to a suitable change
of basis for the ring of polynomials, all coefficients will actually be in
$\Nat$. This relatively simple trick works by considering as ``basic
polynomials" the \intro{binomial function} of the form $\binom{X - \ell}{k}$,
where $\ell$ and $k$ are natural numbers. We extend this to \intro{natural
binomial functions} that are obtained by considering $\Nat$-linear combinations
of products of \kl{binomial functions}, that is, we consider functions that
have the following shape: $f(x_1, \dots, x_k) = \sum_{i = 1}^n n_i \prod_{j =
1}^k \binom{x_j - p_{i,j}}{k_{i,j}}$. Beware that $\binom{X - \ell}{k}$ is
defined to be $0$ when $X \leq \ell$, and is therefore not a polynomial. For
this reason, we also introduce its ``polynomial counterpart" $\intro*\pbinom{X
- \ell}{k}$ defined as $(X - \ell) \cdots (X - \ell - k) / k!$, that we call a
\intro{binomial monomial}, and introduce \intro{natural binomial polynomials}
as $\Nat$-linear combinations of products of \kl{binomial monomials}, i.e., of
the shape: $P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \pbinom{X_j
- p_{i,j}}{k_{i,j}}$.

\begin{lemma}
    \label{binomial-function-star-free:lem}
    Let $F$ be a \kl{binomial function} from $\Nat^k$ to $\Nat$.
    Then it is \kl{represented} by a \kl{star-free polyregular function}.
\end{lemma}
\begin{proof}
    Because of the stability properties of 
    $\NSF$ (\cref{stability-polyregular:lemma}), we only need to 
    check that given $r,s \in \Nat$,
    the function $x \mapsto \binom{x - r}{s}$ is \kl{represented}
    by a function $f_{r,s} \in \NSF$.
    Let us prove it when $r = 0$, since the other functions 
    can be obtained by translating $f_{r,s}$.
    Remark that 
    $\binom{x}{s} = \sum_{u_1 \cdots u_{s+1} = a^x} 1$,
    which is clearly expressible as a \kl{star-free polyregular function}.
\end{proof}

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be an \kl{$\Nat$-polyregular} function. 
    There exists a computable $\omega \in \Nat_{\geq 1}$
    such that for all \kl{polynomial functions}
    $q \colon \Nat^p \to \Sigma^*$,
    there exists a computable \kl{natural binomial function} $F$
    such that:
    \begin{equation*}
        f \circ q(\omega X_1, \dots, \omega X_p)
        =
        F
        \quad 
        \text{ over } (\Nat_{\geq 1})^p
        \quad .
    \end{equation*}
\end{lemma}
\begin{proof}
    Let us fix a finite monoid $M$, a morphism $\mu \colon \Sigma^* \to M$, and 
    let $\omega$ be an idempotent power for the finite monoid $M$,
    i.e., a number such that for all $x \in M$,
    $(x^{\omega})^2 = x^{\omega}$.
    We will prove by induction on $k \in \Nat_{\geq 1}$
    that for all
    $\pi \colon M^k \to \Nat$,
    the lemma holds with $s = 0$ for the function 
    $f = \pi^\dagger$, i.e., such that
    $f(w)$ is the sum over all factorizations of $w$
    into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
    of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.
    Since every \kl{$\Nat$-polyregular function}
    is obtained via some choice of $M, \mu, \pi$ (\cref{nat-rel-poly:def})
    we derive the desired result.

@@

    \textbf{Base case: $k = 1$.}
    Let us show that $f \circ q$ is
    actually a constant (positive) function (and in particular a \kl{natural binomial polynomial}), as for
    all $n_1, \dots, n_p \geq 1$,
    \begin{align*}
        f(q(n_1, \dots, n_p))
        &\eqdef \pi\left(\mu(q(n_1, \dots, n_p))\right) \\
        &\eqdef \pi\left(\mu(\alpha_0 \prod_{i = 1}^p u_i^{\omega n_i} \alpha_i)\right)
        \\
        &= \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i)^{\omega n_i} \mu(\alpha_i)\right) 
        & \text{ morphism }
        \\
        &= \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i) \mu(\alpha_i)\right)
        & \text{ idempotent and } n_i \geq 1 \\
        &= c \geq 0
    \end{align*}

@@

    We will use the fact that given $n_1, \dots, n_p \geq 0$,
    there exists a bijection between
    partitions of $q(\omega n_1, \dots, \omega n_p)$
    into $k+1$ words, and 
    triples composed of
    a position  $1 \leq \ell \leq p$,
    a ratio     $0 \leq b \leq n_\ell$,
    a remainder $0 \leq r < \omega \times \card{u_\ell}$,
    and a partition of
    $)u_\ell^\omega)_{[-r:]} u_\ell^{\omega \times (n_\ell - b)}
    \alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{n_p} \alpha_{p}$
    into $k$ words,
    where we have used the notation 
    $(v)_{[-r:]}$ to denote the word obtained from $v$ 
    by keeping only the last $r$ letters, in a sort of Pythonic syntax.
    That is, we look at where the first ``cut" is made.

    Now, given $(\ell, b, r)$ respecting the inequalities 
    mentioned above,
    one can define the \kl{polynomial functions}
    $s_{\ell,r,b} \colon \Nat^{\ell - 1} \to \Sigma^*$
    and
    $q_{\ell,r} \colon \Nat^{p - \ell + 1} \to \Sigma^*$
    so that
    $q(X_1, \dots, X_p) = s_{\ell,r,b}(X_1, \dots, X_{\ell-1}) 
    q_{\ell, r}(X_{\ell} - b, \dots, X_p)$
    under the assumption that $X_\ell \geq b$:
    \begin{equation*}
        q_{\ell,r}(X_{1}, \dots, X_{p - \ell + 1})
        \defined
        u_\ell^\omega\mid_{[-r:]} u_\ell^{\omega \times X_1}
        \alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{X_{p - \ell + 1}} \alpha_{p}
        \quad .
    \end{equation*}
    Similarly, let us define 
    $s_{\ell, r, b}(X_1, \dots, X_{\ell - 1})$
    as 
    $\alpha_0 u_1^{\omega X_1} \alpha_2 \dots \alpha_{\ell - 1} 
    u_\ell^{\omega b} (u_{\ell}^\omega)_{[1:r]}$,
    Remark that $q_{\ell,r}$ has all of its indeterminates
    multiplied by $\omega$, so that the induction hypothesis 
    can be used.

    The key remark is that $\mu(s_{\ell,r,b}(X_1, \dots, X_{\ell - 1}))$ is a
    constant function from $\Nat^{\ell-1}$ to $M$, and that its value
    only depends on whether $b = 0$ or $b > 0$.
    This holds with
    the exact same proof as the base case of this induction. Let us call
    $m_{\ell,r,1}$ and $m_{\ell,r,0}$ the two elements of the monoid that are
    obtained by $\mu(s_{\ell,r,b}(X_1, \dots, X_{\ell-1}))$ respectively when
    $b = 0$ and when $b > 0$. Finally, let us define $\pi_{\ell,r,0} \colon M^k
    \to \Nat$ via $\pi_{\ell,r,0}(m_2, \dots, m_{k+1}) \defined
    \pi(m_{\ell,r,0}, m_1, \dots, m_{k+1})$.

    With all of this preliminary work done, we are now ready to conclude by
    simply unrolling the definition of 
    a function
    $f$ defined using a \kl{production function}
    $\pi \colon M^{k+1} \to \Nat$.
    In the following equations, we omit the precise
    range for the sum over $(\ell, r, b)$ to save space.
    Similarly, we will use 
    $q_{\ell,r}(X_1, \dots, X_{p-\ell+1})
    \defined q_{\ell,r}(X_1+b, X_2, \dots, X_{p - \ell + 1})$
    at the end of the computation.
    \begin{align*}
        &f(q(n_1, \dots, n_p))\\
        &\eqdef 
        \sum_{v_1 \dots v_{k+1} = q(n_1, \dots, n_p)}
        \pi(\mu(v_1), \dots, \mu(v_{k+1})) \\
        &= \sum_{(\ell, b, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell - b, \dots, n_p)}
           \pi(\mu(s_{\ell,r,b}(n_1, \dots, n_{\ell-1})), \mu(v_2), \dots, \mu(v_{k+1}))
        \\
        &= 
           \sum_{(\ell, 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,b,0}(n_\ell - b, \dots, n_p)}
           \pi(m_{\ell,r,0}, \mu(v_2), \dots, \mu(v_{k+1}))
           \\
        &+ 
           \sum_{(\ell, b > 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell - b, \dots, n_p)}
           \pi(m_{\ell,r,1}, \mu(v_2), \dots, \mu(v_{k+1}))
        \\
        &= 
           \sum_{(\ell, 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,b,0}(n_\ell - b, \dots, n_p)}
           \pi_{\ell,r,0}(\mu(v_2), \dots, \mu(v_{k+1}))
           \\
        &+ 
           \sum_{(\ell, b > 0, r)}
           \sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell-b, \dots, n_p)}
           \pi_{\ell,r,1}(\mu(v_2), \dots, \mu(v_{k+1}))
        \\
        &= 
           \sum_{1 \leq \ell \leq p}
           \sum_{0 \leq r < \omega \times \card{u_\ell}}
           (\pi_{\ell,r,0})^\dagger (q_{\ell, 0, r}(n_\ell-b, \dots, n_p))
           \\
        &+ 
           \sum_{1 \leq \ell \leq p}
           \sum_{0 \leq r < \omega \times \card{u_\ell}}
           \sum_{0 < b < n_\ell}
           (\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
    \end{align*}
    Now, using the induction hypothesis,
    we have \kl{natural binomial functions} $F_{\ell,r,0}$ and $F_{\ell,r,1}$
    satisfying 
    \begin{equation*}
       (\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
       = F_{\ell,r,1}(\omega (n_\ell - b), \dots, \omega n_p)
       \quad 
       \forall n_\ell - b, \dots, n_p \geq 1
       \quad .
   \end{equation*}
   And similarly
    \begin{equation*}
        (\pi_{\ell,r,0})^{\dagger}(q_{\ell,r}(n_1, \dots, n_{\ell-1})
        = F_{\ell,r,0}(\omega n_1, \dots, \omega n_{\ell-1})
       \quad 
       \forall n_1, \dots, n_{\ell - 1} \geq 1
       \quad .
   \end{equation*}

   It is a well-known equality that for all $s,n \in \Nat$,
   $\sum_{0 < i < n} \binom{s}{i} = \binom{s+1}{n}$.
   As a consequence, \kl{natural binomial functions} are
   stable under the following summation operation:
   $\Sigma_Y \colon Q(x_1,\dots,x_n,y) \mapsto \sum_{0 < b < y} Q(x_1, \dots, x_n, b)$.
   This allows us to conclude by defining:
   \begin{equation*}
       F(x_1, \dots, x_p) \defined
       \sum_{1 \leq \ell \leq p} \sum_{0 \leq r < \omega \card{u_\ell}}
       F_{l,r,0}(x_{\ell}, \dots, \omega x_p) + \Sigma_Y F_{l,r,1}(Y,
       x_{\ell+1}, \dots, x_p)
       \quad .
   \end{equation*}
   The latter being a \kl{natural binomial function} satisfying
   the desired equality.

\end{proof}

Let us highlight that the multiplicative factor $\omega$ is necessary in
\cref{n-poly-combinatorics:lem}. Indeed, the function $f \colon \set{a}^* \to
\Nat$ defined as $0$ when the input is of odd length and $1$ when the input is
of even length is \kl{$\Nat$-polyregular}, but $f(a^X)$ is not a polynomial.
We can trade-off this multiplicative factor for a constant term 
addition under the extra assumption that the function 
is \kl{star-free polyregular}, as described in the following lemma.

\begin{lemma}
    \label{n-sf-combinatorics:lem}
    Let $f$ be a \kl{star-free $\Nat$-polyregular} function. 
    There exists a computable $s \in \Nat_{\geq 1}$
    such that for all \kl{polynomial functions}
    $q \colon \Nat^p \to \Sigma^*$,
    there exists a computable \kl{natural binomial function} $F$
    such that:
    \begin{equation*}
        f \circ q(X_1 + s, \dots, X_p + s)
        =
        F
        \quad 
        \text{ over } (\Nat_{\geq 1})^p
        \quad .
    \end{equation*}
\end{lemma}
\begin{proof}[Proof Sketch]
    We redo the proof of \cref{n-poly-combinatorics:lem}, with the extra
    assumption that the monoid $M$ is \kl(monoid){aperiodic}, hence that there
    exists $s \in \Nat$ such that $x^{s+1} = x^s$ for all $x \in M$. As a
    consequence, we can define $\omega = 1$, and
    translate inputs by $s$, instead of multiplying them by $\omega$.
\end{proof}

Let us also address the fact that the distinction between \kl{natural binomial
functions} and \kl{natural binomial polynomials} vanishes when we are
interested in polynomials. As a consequence, to a \kl{natural binomial
function} corresponds exactly one \kl{natural binomial polynomial} that
coincides with it on sufficiently large inputs.

\begin{lemma}
    \label{natural-binomial-polynomial:lem}
    Let $P \in \Rat[X_1, \dots, X_k]$ be a polynomial,
    $\ell \in \Nat$, 
    and let $f$ 
    be a 
    \kl{natural binomial function} such that
    $P = f$ over $(\Nat_{\geq \ell})^n$. If
    $f(x_1, \dots, x_k) = \sum_{i = 1}^n n_i 
    \prod_{j = 1}^k \binom{x_j - p_{i,j}}{k_{i,j}}$,
    then
    $P =
    \sum_{i = 1}^n n_i 
    \prod_{j = 1}^k \pbinom{X_j - p_{i,j}}{k_{i,j}}$.
\end{lemma}
\begin{proof}
    Let $m$ be the maximal value of all $p_{i,j}$, $k_{i,j}$, and $\ell$.
    Over $(\Nat_{\geq 2m})^k$, $f$ coincides with
    the desired polynomial. Now, because $f$,$P$, and the desired
    polynomial coincide
    over a large enough subset of $\Nat^k$,
    this means that $P$ has the desired shape.
\end{proof}

\begin{corollary}
    \label{n-rat-correct:lem}
    Let $P \in \Rel[X_1, \dots, X_p]$ be an \kl{$\Nat$-rational polynomial}.
    Then,
    $P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
    Let $f$ be the \kl{commutative}
    \kl{$\Nat$-polyregular function}
    with domain defined as $\Sigma \defined \set{a_1, \dots, a_p}$
    that \kl{represents} $P$. 
    Using \cref{n-poly-combinatorics:lem},
    there exists a number $\omega \in \Nat_{\geq 1}$
    and \kl{natural binomial function} $Q$
    such that
    for all $n_1, \dots, n_p \geq 1$:
    \begin{equation*}
        f\left(
            \prod_{i = 1}^p a_i^{\omega n_i}
        \right)
        = Q(n_1, \dots, n_p)
        = P(\omega n_1, \dots, \omega n_p) 
        \quad .
    \end{equation*}
    As a consequence of \cref{natural-binomial-polynomial:lem}, $P(\omega X_1, \dots, \omega X_p)$ equals the corresponding 
    \kl{natural binomial polynomial}.
    Let us remark that the (unique) \kl{maximal monomial} of the 
    \kl{binomial monomial} $\pbinom{X -
    p}{k}$ has coefficient $1/k!$, and is therefore \kl{non-negative}.
    We conclude that
    the \kl{maximal monomials} of 
    $P(\omega X_1, \dots, \omega X_p)$ are \kl{non-negative},
    and since $\omega \geq 1$, we conclude that
    the \kl{maximal monomials} of $P$ have \kl{non-negative} coefficients.
